Push-Pull Signal-Function Functional Reactive Programming

### 2012-11-28 Docs/IFL2012/design.tex
Thesis draft updates

The representation of signal functions is discussed in
Section~\ref{section:Implementation-Signal_Functions}.
\begin{figure}
\begin{code}
-- Signal functions
-- init: The initialization type for 
--   the signal function, always NonInitialized
--   for exported signal functions
-- svIn: The input signal vector
-- svOut: The output signal vector
data SF init svIn svOut

@@

The basic signal functions (Figure~\ref{figure:basic_signal_functions})
provide very simple operations. The {\tt identity} signal function, as expected,
simply copies its input to its output. The {\tt constant} signal function
produces the provided value as a signal at all times. The {\tt never} signal
function has an event output which never produces occurrences. The {\tt asap}
function produces an event occurrence with the given value at the first time
step after it is switched into the network. The {\tt after} function waits for
the specified amount of time before producing an event occurrence.

@@

Reactivity is introduced by means of the {\tt switch} combinator
(Figure~\ref{figure:switch_combinator}). The design of this combinator
allows modular packaging of reactivity. A signal function can determine
autonomously when to replace itself, based only on its input and state,
by emitting an event occurrence carrying its replacement. The combinator
consumes and hides the event carrying the replacement signal function, 
so the reactivity is not exposed by the resulting reactive signal function.

It is often useful for a signal function to receive a portion of its
own output as input. This is especially useful when we have two
signal functions which we would like to mutually interact. We cannot
just serially compose them, we must also bring the output of the second
back around to the first. Many signal-processing algorithms also depend
on feedback. The combinator which provides this ability is shown
in Figure~\ref{figure:feedback_combinator}.

@@

The accumulator is useful when responses to multiple event occurrences
(from one or more sources) must be coordinated. For instance, in the
benchmark application (see Chapter~\ref{chapter:Evaluation_and_Comparisons}),
a table is maintained that allows knowledge from previous event occurrences
(packets from a network switch) to be used in deciding where the present
packet ought to go.

The joining combinators provide the ability to combine two event
streams, two signals, or a signal and an event stream. These
combinators are shown in Figure~\ref{figure:joining_combinators}

@@

A set of combinators are provided for making use of time-dependence in a signal
function. These combinators allow the modification of signals and events with
respect to time, and the observation of the current time value. The combinators
are shown in Figure~\ref{figure:time_combinators}

@@

The evaluation interface provides a modified state monad which holds a signal
function, together with some additional information, as its state (shown in Figure~\ref{figure:evaluation_state}.
Rather than monadic instructions to put and get the state, the monad provides instructions
to trigger an input event, update an input signal, and trigger sampling of
signals in the signal function. Additional state includes the current set of
modifications to the input signals (since the last sample) and a set of
handlers which actuate effects based on output events or changes to the output
signal.

@@

In order to initialize the state, the user must supply a set of handlers, the
signal function to evaluate, and initial values for all of the signal inputs
(Figure~\ref{figure:initial_input}).

This state can then be passed to a monadic action which will supply input to
the signal function. Inputs are constructed using a simple interface with
functions to construct sample updates and event occurrences, and to specify
their place in the vector (Figure~\ref{figure:ongoing_input}).

@@

{\tt SFEvalT} actions are constructed using combinators to push events,
update inputs, and step time, as well as actions lifted from the underlying
monad (used to obtain these inputs). An action in the underlying monad
which produces the result and a new state is obtained with the {\tt runSFEvalT}
function. These combinators are shown in Figure~\ref{figure:evaluation_combinators}.

>>>>

The representation of signal functions is discussed in
Section~\ref{subsection:Implementation-Signal_Functions}.
\begin{figure}
\begin{code}
-- Signal functions
-- init: The initialization type for 
--   the signal function, always NonInitialized
--   for exported signal functions
-- svIn: The input signal vector
-- svOut: The output signal vector
data SF init svIn svOut

@@

The basic signal functions
provide very simple operations. The {\tt identity} signal function, as expected,
simply copies its input to its output. The {\tt constant} signal function
produces the provided value as a signal at all times. The {\tt never} signal
function has an event output which never produces occurrences. The {\tt asap}
function produces an event occurrence with the given value at the first time
step after it is switched into the network. The {\tt after} function waits for
the specified amount of time before producing an event occurrence.

@@

Reactivity is introduced by means of the {\tt switch} combinator.
The design of this combinator
allows modular packaging of reactivity. A signal function can determine
autonomously when to replace itself, based only on its input and state,
by emitting an event occurrence carrying its replacement. The combinator
consumes and hides the event carrying the replacement signal function, 
so the reactivity is not exposed by the resulting reactive signal function.

It is often useful for a signal function to receive a portion of its
own output as input. This is especially useful when we have two
signal functions which we would like to mutually interact. We cannot
just serially compose them, we must also bring the output of the second
back around to the first. Many signal-processing algorithms also depend
on feedback. This ability is provided by the {\tt loop} operator.

@@

The joining combinators provide the ability to combine two event
streams, two signals, or a signal and an event stream.

@@

A set of combinators are provided for making use of time-dependence in a signal
function. These combinators allow the modification of signals and events with
respect to time, and the observation of the current time value.

@@

The evaluation interface provides a modified state monad which holds a signal
function, together with some additional information, as its state.
Rather than monadic instructions to put and get the state, the monad provides instructions
to trigger an input event, update an input signal, and trigger sampling of
signals in the signal function. Additional state includes the current set of
modifications to the input signals (since the last sample) and a set of
handlers which actuate effects based on output events or changes to the output
signal.

@@

In order to initialize the state, the user must supply a set of handlers, the
signal function to evaluate, and initial values for all of the signal inputs.

This state can then be passed to a monadic action which will supply input to
the signal function. Inputs are constructed using a simple interface with
functions to construct sample updates and event occurrences, and to specify
their place in the vector.

@@

{\tt SFEvalT} actions are constructed using combinators to push events,
update inputs, and step time, as well as actions lifted from the underlying
monad (used to obtain these inputs). An action in the underlying monad
which produces the result and a new state is obtained with the {\tt runSFEvalT}
function.
